---
author: Abigail I. Gutierrez 
title: "Mini-exercise 3"
output: html_notebook
---

## Mini-exercise 3

```{r}
library(labdsv)
library(plyr)
```


After using all the code above and perusing results, review what you've done with the PICANTE package by going back and making comparisons with the VEGAN package, which doesn't ever consider phylogeny. Simply work with the fake communities created: one, three, two and five.

Then, leave fake communities behind! Do one of the following. You will need to write proper code and show results (with annotations) of your work in this notebook or create a new notebook: 

A. Based on six different recipes or menu items, create lists of at least 10 types of plants that are ingredients. For each, list the 10 most critical plant species that you would need to purchase as ingredients to prepare these foods. For simplicity, list the same number of plant species for each. 

B. Different municipal areas maintain street trees as part of the urban environment. Many publish lists of the most frequent street trees; suggestions below. For six different municipalities, get a list of trees. Have the same number of trees on each list (e.g., ten species each).  (Resources below to give a head-start with this)

Compile your six lists into an abundance matrix. If it's too hard to get abundances, compile presence/absence matrix. Be sure to write the file and push it back to the GitHub repository when you complete this assignment (revisit Digital Lab 2 skills). 

The compiled matrix has a master list of all plant species across all six lists. Resolve this master list for proper taxonomy -- each plant species should have an accepted scientific species name. (Using taxize if possible.)

Phylogenize (revisiting skills from Digital Lab 3). Again, it's a good idea to write the tree you create as a file; you then can and should push back to GitHub as you submit your assignment. 

Create six different annotations of the phylogeny, differing by showing the list-specific taxa. 
Figure out both Shannon or Simpson diversity indices, and PD (Phylogenetic Diversity). Comment on how they compare. Figure out one additional analysis from today's lab, your choice. 

#Part 1. Creating six different lists: Recipies 
I was able to use NYC, City-wide 
http://media.nycgovparks.org/images/web/TreesCount/Index.html to help me get the top 10 species for each borough 
#Part 2. Getting abundance matrixes
```{r}
#Bronx Top 10 Species 
Bronx=read.csv("Bronx.csv")
Bronx
head(Bronx)
tail(Bronx) 

Bronx_list=count(Bronx)
Bronx_list <- subset(Bronx_list[,2:3])
head(Bronx_list)
Bronx_list=data.frame(Bronx_list)
Bronx_list=data.frame("Bronx", Bronx_list)
head(Bronx_list) 
names(Bronx_list)<-c("ListID", "SpeciesID", "Freq") 

#Brooklyn Top 10 Species 
Brooklyn=read.csv("Brooklyn.csv")
Brooklyn
head(Brooklyn)
tail(Brooklyn) 


Brooklyn_list=count(Brooklyn)  
Brooklyn_list <- subset(Brooklyn_list[,2:3]) 
head(Brooklyn_list)
Brooklyn_list=data.frame(Brooklyn_list)
Brooklyn_list=data.frame("Brooklyn", Brooklyn_list)
head(Brooklyn_list) 
names(Brooklyn_list)<-c("ListID", "SpeciesID", "Freq") 

#Staten Island Top 10 Species
StatenIsland=read.csv("StatenI.csv")
StatenIsland
head(StatenIsland)
tail(StatenIsland) 

StatenIsland_list=count(StatenIsland) 
StatenIsland_list <- subset(StatenIsland_list[,2:3]) 
head(StatenIsland_list)
StatenIsland_list=data.frame(StatenIsland_list)
StatenIsland_list=data.frame("Staten Island", StatenIsland_list)
head(StatenIsland_list) 
names(StatenIsland_list)<-c("ListID", "SpeciesID", "Freq") 


#Queens Top 10 Species
Queens=read.csv("Queens.csv")
Queens
head(Queens)
tail(Queens) 

Queens_list=count(Queens)
Queens_list <- subset(Queens_list[,2:3])  
head(Queens_list)
Queens_list=data.frame(Queens_list)
Queens_list=data.frame("Queens", Queens_list)
head(Queens_list) 
names(Queens_list)<-c("ListID", "SpeciesID", "Freq") 

#Manhattan Top 10 Species
Manhattan=read.csv("Manhattan.csv")
Manhattan
head(Manhattan)
tail(Manhattan) 

Manhattan_list=count(Manhattan)
Manhattan_list <- subset(Manhattan_list[,2:3]) 
head(Manhattan_list)
Manhattan_list=data.frame(Manhattan_list)
Manhattan_list=data.frame("Manhattan", Manhattan_list)
head(Manhattan_list) 
names(Manhattan_list)<-c("ListID", "SpeciesID", "Freq") 

#City Top 10 Species 
City=read.csv("City.csv")
City
head(City)
tail(City) 


City_list=count(City)
City_list <- subset(City_list[,2:3]) 
head(City_list)
City_list=data.frame(City_list)
City_list=data.frame("City", City_list)
head(City_list) 
names(City_list)<-c("ListID", "SpeciesID", "Freq") 

```

#Part 3. Master List & Phylogenize 
```{r}
#Combining all data sets into one 
Top10 <- rbind(City_list, Queens_list, Manhattan_list, Brooklyn_list, Bronx_list, StatenIsland_list)
dim(Top10)

#run the matrify function 
CityTree=matrify(Top10)
CityTree 


```

#Part 4: Annotations & Shannon or Simpson diversity indices 

```{r}
abundance <- as.matrix(CityTree)
dim(abundance) 

plot(CityTree)
CityTree
prunedAB <- prune.sample(abundance, CityTree)
plot(prunedAB)

par(mfrow = c(2, 2)) 
for (i in row.names(abundance)) {
plot(prunedAB, show.tip.label = FALSE, main = i)
  tiplabels(tip = which(prunedAB$tip.label %in% names(which(abundance [i, ] > 0))), pch=19, cex=1)
} 

prunedAB
View(prunedAB)

library(vegan)
simpson<-diversity(CityTree, index = "simpson")
shannon<-diversity(CityTree, index = "shannon")
simpson
shannon
jaccard = vegdist(CityTree, "jaccard")
jaccard
bray = vegdist(CityTree, "bray")
bray


```
Somehow, I was able to get the simpson and shannon index to work. Codes 150 -167 now work. I was able to find that the simpson diversity was a 0.9 for each of the 6 boroughs I looked at this meant that 
they were all equal in terms of diversity indexes. In this portfolio I decided to look at the most popular trees in NYC.It makes sense that the shannon indexes all equal to each other just because I noticed that all of the species I was looking at were mostly the same except for a few here and there. The simpson value outputed was 0.9 and this meant that the different boroughs were equal in evenness and in abundance. The Shannon divesity output also gave me the same value for all of them which was 2.3025. This value determined the diversity of the different boroughs. It would be interesting if there was some way to input all of the plant species and account for the diversities of the different boroughs. 
The most interesting measure of diversity became visible in the Jaccard index. This measure revealed the similarity beterrn Manhattan and Staten Island at 0.75. The lowest measure was between Brooklyn and the Bronx and the City's most popular trees. 
The Bray Index showed a high index between Manhattan and Queens with a value of 0.4 but with a much higher value between Manhattan and Staten Island. 

```{r}
taxa<-c( "Acer", "Acer rubum", "Acer platanoides", "Pyrus calleryana", "Prunus cerasus", "Gleditsia triacanthos", "Platanus × acerifolia","Quercus palustris", "Sophora", "Tilia cordata", "Prunus cerasus", "Zelkova serrata") 

phylomatic <- phylomatic(taxa = taxa, storedtree ="R20120829", get = 'POST')

plot(phylomatic, no.margin = TRUE) 

newick_fn = "phylomatic"
write.tree(phylomatic, file=newick_fn)
```
After plotting the Phylogenetic Tree I saw that the outgroup is the platanus aceriflora which is the London PlaneTree. 





