---
author: Abigail I. Gutierrez 
title: "Mini-exercise 3"
output: html_notebook
---

## Mini-exercise 3

After using all the code above and perusing results, review what you've done with the PICANTE package by going back and making comparisons with the VEGAN package, which doesn't ever consider phylogeny. Simply work with the fake communities created: one, three, two and five.

Then, leave fake communities behind! Do one of the following. You will need to write proper code and show results (with annotations) of your work in this notebook or create a new notebook: 

A. Based on six different recipes or menu items, create lists of at least 10 types of plants that are ingredients. For each, list the 10 most critical plant species that you would need to purchase as ingredients to prepare these foods. For simplicity, list the same number of plant species for each. 

B. Different municipal areas maintain street trees as part of the urban environment. Many publish lists of the most frequent street trees; suggestions below. For six different municipalities, get a list of trees. Have the same number of trees on each list (e.g., ten species each).  (Resources below to give a head-start with this)

Compile your six lists into an abundance matrix. If it's too hard to get abundances, compile presence/absence matrix. Be sure to write the file and push it back to the GitHub repository when you complete this assignment (revisit Digital Lab 2 skills). 

The compiled matrix has a master list of all plant species across all six lists. Resolve this master list for proper taxonomy -- each plant species should have an accepted scientific species name. (Using taxize if possible.)

Phylogenize (revisiting skills from Digital Lab 3). Again, it's a good idea to write the tree you create as a file; you then can and should push back to GitHub as you submit your assignment. 

Create six different annotations of the phylogeny, differing by showing the list-specific taxa. 
Figure out both Shannon or Simpson diversity indices, and PD (Phylogenetic Diversity). Comment on how they compare. Figure out one additional analysis from today's lab, your choice. 

#Part 1. Creating six different lists: Recipies 
I was able to use NYC, City-wide 
http://media.nycgovparks.org/images/web/TreesCount/Index.html to help me get the top 10 species for each borough 
#Part 2. Getting abundance matrixes
```{r}
#Bronx Top 10 Species 
Bronx=read.csv("Bronx.csv")
Bronx
head(Bronx)
tail(Bronx) 
speciesID1=paste(Bronx$Common.Name, Bronx$Scientific.Name, sep=" ")
speciesID1

Bronx_list=count(Bronx)
head(Bronx_list)
Bronx_list=data.frame(Bronx_list)
Bronx_list=data.frame("Bronx", Bronx_list)
head(Bronx_list) 
names(Bronx_list)<-c("ListID", "SpeciesID1", "Freq") 

#Brooklyn Top 10 Species 
Brooklyn=read.csv("Brooklyn.csv")
Brooklyn
head(Brooklyn)
tail(Brooklyn) 
speciesID2=paste(Brooklyn$Common.Name, Brooklyn$Scientific.Name, sep=" ")
speciesID2
Brooklyn_list=count(Brooklyn)  

head(Brooklyn_list)
Brooklyn_list=data.frame(Brooklyn_list)
Brooklyn_list=data.frame("Brooklyn", Brooklyn_list)
head(Brooklyn_list) 
names(Brooklyn_list)<-c("ListID", "SpeciesID2", "Freq") 

#Staten Island Top 10 Species
StatenIsland=read.csv("StatenI.csv")
StatenIsland
head(StatenIsland)
tail(StatenIsland) 
speciesID3=paste(StatenIsland$Common.Name, StatenIsland$Scientific.Name, sep=" ")
speciesID3
StatenIsland_list=count(StatenIsland) 

head(StatenIsland_list)
StatenIsland_list=data.frame(StatenIsland_list)
StatenIsland_list=data.frame("Staten Island", StatenIsland_list)
head(StatenIsland_list) 
names(StatenIsland_list)<-c("ListID", "SpeciesID3", "Freq") 


#Queens Top 10 Species
Queens=read.csv("Queens.csv")
Queens
head(Queens)
tail(Queens) 
speciesID4=paste(Queens$Common.Name, Queens$Scientific.Name, sep=" ")
speciesID4
Queens_list=count(Queens)

head(Queens_list)
Queens_list=data.frame(Queens_list)
Queens_list=data.frame("Queens", Queens_list)
head(Queens_list) 
names(Queens_list)<-c("ListID", "SpeciesID4", "Freq") 

#Manhattan Top 10 Species
Manhattan=read.csv("Manhattan.csv")
Manhattan
head(Manhattan)
tail(Manhattan) 
speciesID5=paste(Manhattan$Common.Name, Manhattan$Scientific.Name, sep=" ")
speciesID5
Manhattan_list=count(Manhattan)

head(Manhattan_list)
Manhattan_list=data.frame(Manhattan_list)
Manhattan_list=data.frame("Manhattan", Manhattan_list )
head(Manhattan_list) 
names(Manhattan_list)<-c("ListID", "SpeciesID5", "Freq") 

#City Top 10 Species 
City=read.csv("City.csv")
City
head(City)
tail(City) 
speciesID6=paste(City$Common.Name, City$Scientific.Name, sep=" ")
speciesID6
City_list=count(City)

head(City_list)
City_list=data.frame(City_list)
City_list=data.frame("City", City_list )
head(City_list) 
names(City_list)<-c("ListID", "SpeciesID6", "Freq") 

```

#Part 3. Master List & Phylogenize 
```{r}
#Combining all data sets into one 
Top10=rbind(City_list, Queens_list, Manhattan_list, Brooklyn_list, Bronx_list, StatenIsland_list)
dim(Top10) 

#run the matrify function 
CityTree=matrify(Top10)
CityTree 
```

#Part 4: Annotations & Shannon or Simpson diversity indices 
```{r}
simpson2<-diversity(wab1, index = "simpson")
shannon2<-diversity(wab1, index = "shannon")
simpson2
shannon2
jaccard2 = vegdist(wab1, "jaccard")
jaccard2
bray2 = vegdist(wab1, "bray")
bray2  
```

