---
author: Abigail I. Gutierrez 
title: "R Notebook"
output: html_notebook
---

## Mini-exercise #2

Let's build and combine your new skills in looking up species occurrence data with what we learned about mapping. 

Going back to our peninsula, here are a bunch of species that are endemic to that area.

Spergularia melanocaulos
Petrocoptis grandiflora
Spergula viscosa
Minuartia cerastiifolia
Minuartia recurva
Petrocoptis hispanica
Petrocoptis pyrenaica

Choose any two, and do the following:
1. Minuartia cerastiifolia
2. Petrocoptis grandiflora 

WORK WITH TAXISE
1. Look up - are they accepted names or are they synonyms?

Minuartia cerastiifolia is an accepted name for Minuartia cerastiifolia. According to the code I put in below however for the code on the chunk below this I got that the status was SYNONYM. Their species name is Facchinia cerastiifolia. 

Petrocoptis pyrenaica  is an accepted name for Petrocoptis pyrenaica A.Braun ex Walp. 

```{r}
library(BIEN)
library(ape) # Package for working with phylogenies in R
library(maps) # Useful for making quick maps of occurrences
library(sp) # A package for spatial data
library(raster)# Another package for spatial data 
vignette("BIEN") 
results=gnr_resolve(names="Minuartia cerastiifolia")
head(results$matched_name)
results=gnr_resolve(names="Petrocoptis pyrenaica")
head(results$matched_name)

```

2. Look up the two plant taxa. Are they both angiosperms?
Minuartia cerastiifolia is in the phylum of Tracheophyta 

Petrocoptis pyrenaica is in the phylum of Tracheophyta 
```{r}

classification(c("Minuartia cerastiifolia", "Petrocoptis grandiflora"), db="itis") 

#Minuartia cerastiifolia 
name="Minuartia cerastiifolia"  
tsn=get_tsn(name, accepted=FALSE)
tsn 
lapply(tsn, itis_acceptname) # there was an error for this line of code 

name_lookup(q="Minuartia cerastiifolia")
name_backbone(name="Minuartia cerastiifolia")
name_usage(8998109) 

#Here's another way to do that, in another database
tax_name(q="Minuartia cerastiifolia", get="family", db="ncbi") 
#here I got an error because it is reading as a character 

#Petrocoptis pyrenaica
name="Petrocoptis pyrenaica"  
tsn=get_tsn(name, accepted=FALSE)
tsn 
lapply(tsn, itis_acceptname) # there was an error for this line of code 

name_lookup(q="Petrocoptis pyrenaica")
name_backbone(name="Petrocoptis pyrenaica")
name_usage(7267085)  

```

3. Angiosperm or otherwise, what plant family do they belong to?
Minuartia cerastiifolia , Family: Caryophyllaceae 
Petrocoptis pyrenaica, Family: Caryophyllaceae 

```{r}
tax_name(q="Minuartia cerastiifolia", get="family", db="ncbi") 
 
tax_name(q="Petrocoptis grandifloras", get="family", db="ncbi")  


```

4. For at least one genus, look up all the species in that genus.
For Petrocoptis pyrenaica, I looked at the genus of Petrocoptis it was able to retrieve data but got me nothing. For Minuartia I got 37 species that resulted in the genus Minuartia. 
```{r}
#Petrocoptis pyrenaica 
downstream("Petrocoptis", downto="species", db="itis")

#Minuartia cerastiifolia
downstream("Minuartia", downto="species", db="itis")
```


5. For one species, look up to see if there are subspecies or other infraspecific forms.
   the code resulted in 0 rows and 0 length which tell me that there are no subspecies or other infraspecific forms. 
```{r}
children(" Minuartia", db="col") 
```


WORK WITH RGBIF OR SPOCC
6. For both, search and download species occurrence data; prepare as a data-frame

```{r}
#Minuartia cerastiifolia 
Minuartia <- occ(q="Minuartia cerastiifolia", from="gbif", limit=50)
Minuartia

Minuartia.dat <- (Minuartia$gbif$data)
head(Minuartia.dat)

Minuartia.df <- occ2df(Minuartia)
Minuartia.df

dim(Minuartia.df)
Minuartia.df <- Minuartia.df[complete.cases(Minuartia.df ),]
dim(Minuartia.df)
Minuartia.df


#now working with the next data Petrocoptis pyrenaica  
pyrenaica <- occ(q="Petrocoptis pyrenaica", from="gbif", limit=50)
pyrenaica

pyrenaica.dat <- (pyrenaica$gbif$data)
head(pyrenaica.dat)

pyrenaica.df <- occ2df(pyrenaica)
pyrenaica.df

dim(pyrenaica.df)
pyrenaica.df<- pyrenaica.df[complete.cases(pyrenaica.df),]
dim(pyrenaica.df)
pyrenaica.df



```

7. Clean-up the data-frame:  remove occurrences with missing data and duplicates
did that above 

WORK WITH OCCURRENCE DATA, WORLDCLIM DATA AND MAPS
8. Map occurrence data with a climate variable for one species on the peninsula

```{r}
P_pyrenaica <- BIEN_occurrence_species(species = "Petrocoptis pyrenaica") 
P_pyrenaica_rangemap <- BIEN_ranges_load_species("Petrocoptis pyrenaica") 

map('world', fill=T, col="grey", bg="light blue", xlim=c(-10,4), ylim=c(35,45))
  plot(P_pyrenaica_rangemap, col="green", add= T) 

  
all.worldclim <- raster::getData("worldclim", res = 10, var = "bio")
plot(all.worldclim[["bio1"]])
points(P_pyrenaica[,c("longitude", "latitude")], pch = 24, cex = 0.5)

```


9. Also map that species with climate variables on a world map or European map. Are you convinced that the species is endemic to this peninsula? 
```{r}
peninsula.worldclim <- crop(all.worldclim, extent(-10, 4, 35, 45))
plot(peninsula.worldclim[["bio1"]])
```



10. Returning to your results for #4 above, find another species with a range outside of the  peninsula. Download occurrence data and map on a world map. 
  Minuartia macrocarpa 
  
```{r}
P_macrocarpa  <- BIEN_occurrence_species(species = "Minuartia macrocarpa") 
P_macrocarpa_rangemap <- BIEN_ranges_load_species("Minuartia macrocarpa") 
P_macrocarpa_rangemap

map('world', fill=T, col="grey", bg="light blue", xlim=c(-10,4), ylim=c(35,45))
  plot(P_macrocarpa_rangemap, col="green", add= T) 

  
all.worldclim <- raster::getData("worldclim", res = 10, var = "bio")
plot(all.worldclim[["bio1"]])
points(P_macrocarpa[,c("longitude", "latitude")], pch = 24, cex = 0.5)

plot(all.worldclim[["bio1"]])
points(P_macrocarpa[,c("longitude", "latitude")], pch = 24, cex = 0.5)
```


11. Then re-map on an appropriately cropped map.
```{r}
plot(all.worldalt[["alt"]], xlim=c(-150,-100),ylim=c(50,100))
```


ONE LAST THING
12. Don't forget to identify the peninsula that we are working with!
We are working with the Iberian Peninsula.